The Weakly Nonlinear Saturation of the Magnetorotational Instability

### 2014-09-18 paper/mri_paper.tex
more of jeff's comments, some intro

This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.

@@

The practical advantage of our ansatz construction (Equation \ref{V1_ansatz}) is clear: the separable x-dependence means that the radial boundary conditions are solved in only one dimension. Thus our analytical framework is able to side-step many of the resolution issues faced by multidimensional simulations. We are able to resolve even significant structure in the boundary layers of our domain, because we need only resolve it in one dimension. We solve the radial component of each equation using the open source pseudospectral code Dedalus (Burns et al., in prep). We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains (e.g. \citei{Boyd:2001aa}).

@@

To find a bounded solution at each order, we eliminate secular terms -- terms which are resonant with solutions to the homogenous equation, and cause the solution to grow without bound. The elimination of secular terms requires enforcement of solvability criteria, which arise as a consequence of the Fredholm alternative (e.g. Guenther \& Lee 1988). The solvability criterion is found by forcing the inner product of the equation with its adjoint homogenous solution to be zero. For instance, Equation \ref{ordere2} contains a term $\widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ that is resonant with $e^{i k_c z}$. Because the adjoint homogenous equation

@@

We conduct a full study of the MRI parameter space, focussed on the behavior with changing magnetic Prandtl number ($\mathrm{Pm} \equiv \frac{\mathrm{Rm}}{\mathrm{Re}} \equiv \frac{\nu}{\eta}$). In protoplanetary disks, as well as in liquid metal MRI experiments, the magnetic Prandtl number can be very small ($\mathrm{Pm} \sim 10^{-6}$). The scale separations caused by these large ratios of diffusivity to viscosity render the small Prandtl number regime inaccessible to simulations. This analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of [....].

>>>>

This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime. The present work adds to this foundation both in problems analyzed, with our analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space.

@@

The practical advantage of our ansatz construction (Equation \ref{V1_ansatz}) is clear: the separable x-dependence means that the radial boundary conditions are solved in only one dimension. Thus our analytical framework is able to side-step many of the resolution issues faced by multidimensional simulations. We are able to resolve even small-scale structure in the boundary layers of our domain, because we need only resolve it in one dimension. We solve the radial component of each equation using the open source pseudospectral code Dedalus (Burns et al., in prep). We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains (e.g. \citei{Boyd:2001aa}). The nonuniform spacing of the Chebyshev grid allows us to resolve the boundary structures very well on 256-point grid.

@@

To find a bounded solution at each order, we eliminate secular terms: terms which are resonant with solutions to the homogenous equation and cause the solution to grow without bound. The elimination of secular terms requires enforcement of solvability criteria, which arise as a consequence of the Fredholm alternative (e.g. Guenther \& Lee 1988). The solvability criterion is found by forcing the inner product of the equation with its adjoint homogenous solution to be zero. For instance, Equation \ref{ordere2} contains a term $\widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ that is resonant with $e^{i k_c z}$. Because the adjoint homogenous equation

@@

We conduct a full study of the MRI parameter space, focussed on the behavior with changing magnetic Prandtl number ($\mathrm{Pm} \equiv \frac{\mathrm{Rm}}{\mathrm{Re}} \equiv \frac{\nu}{\eta}$). In protoplanetary disks, as well as in liquid metal MRI experiments, the magnetic Prandtl number can be very small ($\mathrm{Pm} \sim 10^{-6}$). The scale separations caused by these large ratios of diffusivity to viscosity render the small Prandtl number regime inaccessible to simulations. This analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of [TBD!].
